The Ranked Nearest Neighbour Interchange space

### 2019-06-27 paper.tex
authors + some definitions trim

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[foot]{amsaddr}
\usepackage[hidelinks]{hyperref}
\usepackage[style=authoryear,ibidtracker=false,uniquename=false,giveninits=true,terseinits=true,maxbibnames=5,backend=biber]{biblatex}
\renewbibmacro{in:}{}
\addbibresource{lit.bib}

@@

\title{The Ranked Nearest Neighbour Interchange space}
\date{\today}
\author{Lena Collienne}
\email{lena.collienne@postgrad.otago.ac.nz}
\author{Alex Gavryushkin}
\email{alex@biods.org}
\address{Department of Computer Science, University of Otago, New Zealand}

@@

Furthermore, we will identify leaves by their labels and might call them \emph{taxa} (singular: \emph{taxon}).
For a tree $T$ a subtree consisting of an internal node with both children being leaves is called a \emph{cherry}.
An edge $e$ that has no leaf as child is an \emph{internal edge} and we say an edge $e=(u,v)$ has \emph{length} $l$ if $|rank(v) - rank(u)| = l$.
Each internal node $v$ gives a \emph{cluster} $C$ that is the set of taxa descending from this node. \todo{OK?}
We then say that $v$ induces the cluster $C$.
Note that all trees on $n$ taxa share the cluster $\{1,\ldots,n\}$ induced by the root.
A \emph{split} is a bipartition of the taxon set $X$ into sets $A,B$ such that there is an edge $e$ whose deletion results in two trees with taxon sets $A$ and $B$, respectively.
Such a split will be denoted by $A|B$ and we say that it is induced by edge $e$.
Given a subset $S$ of the set of taxa $X$, the \emph{most recent common ancestor} $mrca(S)$ of $S$ is the internal node with lowest rank that is ancestor of all taxa in $S$. \todo{OK?}
In particular, for a single taxon $x \in X$, $\mrca(\{x\})$ is the parent of $x$ and will be denoted by $\parent(x)$.
We write $\mrca_T(S)$ to emphasise that we consider the most recent common ancestor of set $S$ in tree $T$.
Two trees are called \emph{isomorphic} if the underlying rooted phylogenetic trees are isomorphic and the isomorphism preserves ranks of the nodes.

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[misc]{ifsym}
\usepackage[foot]{amsaddr}
\usepackage[hidelinks]{hyperref}
\usepackage[style=authoryear,ibidtracker=false,uniquename=false,giveninits=true,terseinits=true,maxbibnames=5,backend=biber]{biblatex}
\renewbibmacro{in:}{}
\addbibresource{lit.bib}

@@

\title{The Ranked Nearest Neighbour Interchange space}
\date{\today}
\author{Lena Collienne\textsuperscript{1}}
\email{lena.collienne@postgrad.otago.ac.nz}
\address{\textsuperscript{1}Department of Computer Science, University of Otago, New Zealand}
\author{Mareike Fischer\textsuperscript{2}}
\email{email@mareikefischer.de}
\address{\textsuperscript{2}Institute for Mathematics and Inofrmatics, University of Greifswald, Germany}
\author{Alex Gavryushkin\textsuperscript{1, \Letter}}
\email{alex@biods.org}

@@

We will identify leaves with their labels and also call them \emph{taxa} (singular: \emph{taxon}) following the systematic biology tradition.
For a tree $T$, a subtree consisting of an internal node with both children being leaves is called a \emph{cherry}.
An edge $e$ that has no leaf as a child is called an \emph{internal edge} and we say an edge $e=(u,v)$ has \emph{length} $l$ if $|rank(v) - rank(u)| = l$.
Each internal node $v$ can be uniquely identified by its \emph{cluster} $C$, which is the set of taxa descending from the node.
Note that all trees on $n$ taxa share the cluster $\{1,\ldots,n\}$ induced by the root.
A \emph{split} is a bipartition of the taxon set $X$ into sets $A,B$ such that there is an edge $e$ whose deletion results in two trees with taxon sets $A$ and $B$, respectively.
Such a split will be denoted by $A|B$ and we say that it is induced by edge $e$.
Given a subset $S$ of the set of taxa $X$, the \emph{most recent common ancestor} $mrca(S)$ of $S$ is the internal node with lowest rank that is ancestor of all taxa in $S$. \todo{OK?}
In particular, for a single taxon $x \in X$, $\mrca(\{x\})$ is the parent of $x$ and will be denoted by $\parent(x)$.
We write $\mrca_T(S)$ to emphasise that we consider the most recent common ancestor of set $S$ in tree $T$.
Two trees are called \emph{isomorphic} if the underlying rooted phylogenetic trees are isomorphic and the isomorphism preserves ranks of the nodes.
